Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory

### 2019-10-15 paper2_JNLS/paper.tex
NM

 %----------------------------------------------------------------------------%
% Skewness Ratio 
%----------------------------------------------------------------------------%
\section{Predicted outgoing skewness and experimental confirmation}

We now discuss in this section arguably the most significant single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Below, we recap the derivation of this formula and then test the prediction against our direct experimental measurements.

\subsection{Explicit formula for outgoing skewness}

@@

	Our next task is to convert this formula to physical variables so that it can be tested against experimental data. First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the relationship, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
\begin{align}
\label{H3convert}
\meandn{\Hthree} = \frac{\pi}{3} \meandn{u^3} = 
\frac{1}{3 \pi^{1/2}} \frac{\meandn{\eta^3}}{ \etastd^3} = 
\frac{1}{3 \pi^{1/2}} \skwdn(\eta)
\end{align}
This equation expresses a directly relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement.
Next, we must convert the $\Htwo$-quantities, which involve surface slope. Here, it is important to remember that, since it is the {\em downstream} Hamiltonian that is matched in \eqref{statmatch}, one must use the {\em downstream} characteristic scale $\lengthscale_+$ to convert back into physical variables. Using the definition of $\Htwo$ in Eq.~\eqref{H3H2} and $\lengthscale_+$ in Eq.~\eqref{scales} gives
\begin{align}
\label{H2convert}
&\meanupdn{\Htwo} = \pi \meanupdn{ u_x^2} = 
\left( \frac{\lamfac \lamdn}{2 \pi \etastd} \right)^2 \varupdn(\eta_x)
\end{align}
Inserting \eqref{H3convert} and \eqref{H2convert} into \eqref{H3H2ratio}, and using the definitions of $C_2$ and $C_3$ from \eqref{C3C2}, gives the remarkably explicit formula
\begin{equation}
\label{SkewPred}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 
%= \frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
\end{equation}
This formula links the skewness of the outgoing wave field to the change in the variance of surface slope. In particular, it predicts that their ratio to scale as the {\em inverse cube of wave amplitude}, $\epsup^{-3}$, and the {\em cube of the depth ratio}, $\dratdn^3$. It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possibly variable combinations, are so intimately related.

 % Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/SkewRat.pdf}
\caption{
Experimental confirmation of the explicit formula for outgoing skewness \eqref{SkewPred}.
(a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes. The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm respectively (green), which are chosen as the representative locations to evaluate \eqref{SkewPred}.
(c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$. For sufficiently large amplitude, the measurements fit closely with the -3 power law predicted by \eqref{SkewPred}.
}
\label{SkewRat}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 %New experimental measurements guided by theory

	Fortunately, it is a relationship that can be tested directly against our experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}. The the prediction draws our attention to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b). For the representative downstream values, $\skwdn(\eta)$ and $\vardn(\eta_x)$, we select the location of the skewness peak, $x=15$ cm downstream of the ADC. We must also select an upstream location for $\varup(\eta_x)$. Fig.~\ref{SkewRat}(b) suggests that an asymptotic value may be reached far upstream of the ADC, but the data does not extend that far. On the other hand, there appears to be a well-defined minimum in $\varup(\eta_x)$ that occurs at the same position for all driving amplitudes. We thus select this minimum position, $x = -9$ cm upstream of the ADC to measure $\varup(\eta_x)$.

>>>>

 %----------------------------------------------------------------------------%
% Skewness Ratio 
%----------------------------------------------------------------------------%
\subsection{Predicted outgoing skewness and experimental confirmation}

This section discusses arguably the most novel single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Below, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.

%\subsection{Explicit formula for outgoing skewness}

@@

	Our next task is to convert this formula to physical variables so that it can be tested against experimental data. First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the definition, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
\begin{align}
\label{H3convert}
\meandn{\Hthree} = \frac{\pi}{3} \meandn{u^3} = 
\frac{1}{3 \pi^{1/2}} \frac{\meandn{\eta^3}}{ \etastd^3} = 
\frac{1}{3 \pi^{1/2}} \skwdn(\eta)
\end{align}
This equation expresses a direct relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement.
Next, we must convert the $\Htwo$-quantities, which involve surface slope. Here, it is important to remember that, since it is the {\em downstream} Hamiltonian that is matched in \eqref{statmatch}, one must use the {\em downstream} characteristic scale $\lengthscale_+$ to convert back into physical variables. Using the definition of $\Htwo$ in Eq.~\eqref{H3H2} and $\lengthscale_+$ in Eq.~\eqref{scales} gives
\begin{align}
\label{H2convert}
&\meanupdn{\Htwo} = \pi \meanupdn{ u_x^2} = 
\left( \frac{\lamfac \lamdn}{2 \pi \etastd} \right)^2 \varupdn(\eta_x)
\end{align}
Inserting \eqref{H3convert} and \eqref{H2convert} into \eqref{H3H2ratio}, and using the definitions of $C_2$ and $C_3$ from \eqref{C3C2}, gives the remarkably explicit formula
\begin{equation}
\label{SkewPred}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 
%= \frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
\end{equation}
This formula links the skewness of the outgoing wave field to the change in the variance of surface slope. In particular, it predicts their ratio to scale as the {\em inverse cube of wave amplitude}, $\epsup^{-3}$, and the {\em cube of the depth ratio}, $\dratdn^3$. It is a rather unexpected relationship, as it would have been difficult to anticipate that the displacement skewness and slope variance, among all possibly variable combinations, are so intimately related.

 % Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/SkewRat.pdf}
\caption{
Experimental confirmation of the explicit formula for outgoing skewness \eqref{SkewPred}.
(a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes. The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm (green) respectively, which are chosen as the representative locations to evaluate \eqref{SkewPred}.
(c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$. For sufficiently large amplitude, the measurements fit closely with the $\epsup^{-3}$ prediction from \eqref{SkewPred}.
}
\label{SkewRat}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 %New experimental measurements guided by theory

	Fortunately, it is a relationship that can be tested directly against the experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}. The the prediction draws focus to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b). To test \eqref{SkewPred}, we must select a representative downstream position to evaluate $\skwdn(\eta)$ and $\vardn(\eta_x)$, and a representative upstream position for $\varup(\eta_x)$. In short, we select the same representative locations as were used to evaluate the histograms in Figs.~\ref{ExpDiagStats}, \ref{uhist}, and \ref{slopehist}, namely $x = 15$ cm downstream and $x = -9$ cm upstream. These positions are indicated by the blue and green vertical dashed lines in Fig.~\ref{SkewRat}(a)--(b). Recall that $x = 15$ cm corresponds to the peak of the downstream skewness and kurtosis. Meanwhile, we see in Fig.~\ref{SkewRat}(b) that $x = -9$ cm corresponds to a slight dip in $\var(\eta_x)$.

	With these locations selected, we evaluate the ratio $\skw(\eta) / \Delta \var(\eta_x)$ for each experiment and plot the result against dimensionless wave amplitude, $\epsup = \etastd / \dup$ on a log-log scale in Fig.~\ref{SkewRat}(c). While the data shows significant variation for small amplitudes, it shows a well defined decreasing trend at larger amplitudes. In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend remarkably well.

 %----------------------------------------------------------------------------%
% Conclusions 
%----------------------------------------------------------------------------%
\section{Concluding remarks}
